---
title: "Homework Exercise 1"
author: "W203 Instructional Team"
date: \today
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

W203 Statistics for Data Science
========================================================
Unit 1 Homework
========================================================


## Exercise

Load the dataset found in the file, cars.csv.

```{r load}
cars = read.csv("/Users/gautamkarnataki/MIDS/summer2018/Stats/Unit1/hw1/cars.csv")
cars
```

1. What are the variables in the file?
```{r}
names(cars)
```


2. Find the mean, median, minimum, maximum, 1st quartile and 3rd quartile for the mpg variable.
```{r}
summary(cars$mpg)
```


3. Create a histogram of the mpg variable. 
```{r}
colors = c("red", "yellow", "green", "violet", "orange", "blue", "pink", "cyan")
hist(cars$mpg, breaks=5, col=colors, main="Histogram of Miles Per Gallon", xlab = "Miles Per Gallon")

# We can increase the number of buckets by increasing the "breaks" and specifying a non-default axis
```


4. What is the standard deviation of mpg variable?
```{r}
sd(cars$mpg)
```


5. What is the variance of mpg variable?
```{r}
var(cars$mpg)
```

6. What is the relationship of the standard deviation to the variance? Why does the standard deviation and variance of the mpg variable differ? 
```{r}
# standard deviation = square_root(variance). 
# To compute variance, first compute the mean, 
# then subtract each value from mean and square each value. 
# Then add all the values and divide by (n-1). 
# To compute SD, get the square root of variance
```


7. How many data points are there for the cyl variable? 
```{r}
# Depends on what we want to do with the variable
# For e.g, if we want to use the variable to compute mean, sd, var etc, 
# we might choose to omit NAs
cyl = cars$cyl
length(cyl[!is.na(cyl)])

# But if we want to use a derivative of the variable - say, (cyl^0), 
# we might still consider NA to be a data point. In which case,
length(cyl)
```

8. What is the mean of the cyl variable?
```{r}
mean(cyl, na.rm = TRUE)
```



